---
title: "ArboMAP: Arbovirus Modeling and Prediction   \nto Forecast Mosquito-Borne Disease Outbreaks"
author: "Summary of Model Outputs (v2.0)   \nJustin K. Davis and Michael C. Wimberly  \n(justinkdavis@ou.edu, mcwimberly@ou.edu)  \nGeography and Environmental Sustainability, University of Oklahoma"
date: "Updated `r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# define some helpful functions
'%!in%' <- function(x,y)!('%in%'(x,y))
round_any = function(x, accuracy, f=round){f(x/accuracy) * accuracy}
options(warn=-1)
```

```{r createfunctions, include=FALSE}
simplifynames <- function(priornames=NULL) {

  # convert to lower case
  priornames <- tolower(priornames)
  
  # remove spaces
  priornames <- gsub(pattern=" ", replacement="", x=priornames, fixed=TRUE)
  
  # remove district and parish
  priornames <- gsub(pattern="county", replacement="", x=priornames, fixed=TRUE)
  priornames <- gsub(pattern="parish", replacement="", x=priornames, fixed=TRUE)

  # return names
  return(priornames)
  
}
```

```{r libraries, include=FALSE}
packages <- c("reshape2", "ggplot2", "gridExtra",
              "lme4","pracma","dplyr","maptools",
              "raster","spdep","mgcv","sp","rgdal",
              "GISTools","data.table","splines","maps",
              "broom","mapproj", "Hmisc", "parallel")
for (package in packages) {
    if (!require(package, character.only=T, quietly=T)) {
        install.packages(package, repos = "http://cran.us.r-project.org")
        library(package, character.only=T)
    }
}
```

```{r setoptions, include=FALSE, echo=FALSE}

# where do we want the outputs?
graphicoutputdir <- ".\\graphical outputs\\"
fullcasematoutputdir <- ".\\case matrix with estimates\\"
mosqmatoutputdir <- ".\\mosquito matrix with estimates\\"

# where are the human data located?
humandatafile <- ".\\human case data\\simulated human case data.csv"

# what is the date of the last human case we're willing to believe?
# probably, cut this off at the end of last year
# DO NOT use any human cases from the year you're modeling
maxobservedhumandate <- as.Date("2017-12-31", "%Y-%m-%d")

# which week are we producing the graphs for?
weekinquestion <- as.Date("2018-08-15", "%Y-%m-%d")

# where are the weather csv files stored?
weatherpathstr <- ".\\weather data\\"
# what is the name of the summary file to be created?
weathersummaryfile <- "weather data summary file.csv"

# which variables do you want to use?
var1name <- "tmeanc"
var2name <- "vpd"

# where are the mosquito test files located?
mosqfile <- ".\\mosquito data\\simulated mosquito tests.csv"

# which district stratification scheme are we using?
stratafile <- ".\\strata\\17-04-20 - classified strata - classic.csv"

# where is the districtshapefile
districtshapefile <- ".\\shapefile\\cb_2014_us_county_5m - in EPSG 5070 - only SD.shp"

# to which two other years do we want to compare the current year's predictions?
compyear1 <- 2012
compyear2 <- 2017

# probably don't want to modify what follows, but you have some options if you're comfortable

# makes sure we round to the previous Sunday, so that this week is included
weekinquestionSun <- weekinquestion - (as.numeric(strftime(weekinquestion, '%u')) %% 7)
weekinquestionSat <- weekinquestionSun + 6
weekinquestionSunstr <- strftime(weekinquestionSun, '%A')
weekinquestionSatstr <- strftime(weekinquestionSat, '%A')

# figure out which year this is and begin all weeks on Sunday
maxmosqyear <- as.numeric(format(weekinquestion, "%Y"))
maxdesiredhumandate <- as.Date(paste(maxmosqyear,
                                     "-12-31",sep=""))
maxdesiredhumandate <- maxdesiredhumandate - (as.numeric(strftime(maxdesiredhumandate, '%u')) %% 7)
maxdesiredhumandatestr <- strftime(maxdesiredhumandate, '%A')

# make sure the max desired human date is no earlier than the week in question
# otherwise, there will be no predictions for the week in question
maxdesiredhumandate <- max(weekinquestionSun+7, maxdesiredhumandate)

# set up lag and regression information
laglen   <- 181
dlagdeg  <- 8
```

# Data used for predictions
## Weather data

```{r weatherdataload, include=FALSE, echo=TRUE}
# load and concat files
weatherlist <- list.files(path=weatherpathstr, pattern="(.csv)", recursive=FALSE)
weather <- data.frame()
for (i in 1:length(weatherlist)) {
  
  if (weatherlist[[i]] != weathersummaryfile) {
  
    tempdf <- read.csv(paste(weatherpathstr, weatherlist[[i]], sep="")) 
    weather <- bind_rows(weather, tempdf)  
    
  }
  
}
weather$date <- as.Date(paste(weather$year,
                              weather$doy,
                              sep="-"),
                        "%Y-%j")

# get rid of duplicated rows
weather$districtdate <- paste(weather$district, weather$date)
weather <- subset(weather, !duplicated(weather$districtdate))

# plot normals and this year
weather <- group_by(weather,
                    doy)
doymet  <- dplyr::summarize(weather,
                     med_tmeanc = quantile(tmeanc, probs=0.50, na.rm=TRUE),
                     med_rmean  = quantile(rmean,  probs=0.50, na.rm=TRUE),
                     med_vpd    = quantile(vpd,    probs=0.50, na.rm=TRUE),
                     max_tmeanc = max(tmeanc, na.rm=TRUE),
                     max_rmean  = max(rmean,  na.rm=TRUE),
                     max_vpd    = max(vpd,    na.rm=TRUE),
                     min_tmeanc = min(tmeanc, na.rm=TRUE),
                     min_rmean  = min(rmean,  na.rm=TRUE),
                     min_vpd    = min(vpd,    na.rm=TRUE))
weather <- ungroup(weather)

thisyear <- max(weather$year, na.rm=TRUE)
thisyear <- subset(weather, year == thisyear)
thisyear <- group_by(thisyear, doy)
thisyear <- dplyr::summarize(thisyear,
                      med_tmeanc = quantile(tmeanc, probs=0.50, na.rm=TRUE),
                      med_rmean  = quantile(rmean,  probs=0.50, na.rm=TRUE),
                      med_vpd    = quantile(vpd,    probs=0.50, na.rm=TRUE),
                      max_tmeanc = max(tmeanc, na.rm=TRUE),
                      max_rmean  = max(rmean,  na.rm=TRUE),
                      max_vpd    = max(vpd,    na.rm=TRUE),
                      min_tmeanc = min(tmeanc, na.rm=TRUE),
                      min_rmean  = min(rmean,  na.rm=TRUE),
                      min_vpd    = min(vpd,    na.rm=TRUE))

tempdf <- left_join(doymet, thisyear, by="doy")
tempdf <- tempdf[!is.na(tempdf$med_tmeanc.x),]

meantemp.x <- round(mean(tempdf$med_tmeanc.x, na.rm=TRUE), 1)
meantemp.y <- round(mean(tempdf$med_tmeanc.y, na.rm=TRUE), 1)
meanvapd.x <- round(mean(tempdf$med_vpd.x, na.rm=TRUE), 1)
meanvapd.y <- round(mean(tempdf$med_vpd.y, na.rm=TRUE), 1)
```

Weather data from the gridMET data set range from `r min(weather$date, na.rm=TRUE)` to `r max(weather$date, na.rm=TRUE)`. Below are graphs of statewide daily averages of `r var1name` and `r var2name`. Observations for the current year are in red. Black is the medium from all other years, and the grey band indicates the max/min ever observed.

```{r weatherplots, fig.width=7, fig.height=5, echo=FALSE}
plot1 <- ggplot() + geom_line(data=doymet, aes(x=doy, y=med_tmeanc)) +
  geom_ribbon(data=doymet, aes(x=doy, ymin=min_tmeanc, ymax=max_tmeanc), alpha=0.3) +
  geom_line(data=thisyear, aes(x=doy, y=med_tmeanc), color="red", size=1) +
  xlab("Day of the year") + ylab(var1name) +
  ggtitle(paste(var1name, 
                max(weather$year, na.rm=TRUE),
                sep=" "))
plot2 <- ggplot() + geom_line(data=doymet, aes(x=doy, y=log(med_vpd+1))) +
  geom_ribbon(data=doymet, aes(x=doy, ymin=log(min_vpd+1), ymax=log(max_vpd+1)), alpha=0.3) +
  geom_line(data=thisyear, aes(x=doy, y=log(med_vpd+1)), color="red", size=1) +
  ggtitle(paste(var2name,
                max(weather$year, na.rm=TRUE),
                sep=" ")) +
  xlab("Day of the year") + ylab(var2name)
grid.arrange(plot1, plot2, nrow=2)

# simplify the district names
weather$district <- simplifynames(weather$district)
```
```{r silentsaveweather, include=FALSE, echo=FALSE}
# put aside this set for use in the regression
dailyextr <- weather
rm(weather)
gc()

ggsave(paste(graphicoutputdir,
             "var1.png",
             sep=""), plot1)
ggsave(paste(graphicoutputdir,
             "var2.png",
             sep=""), plot2)
```

\newpage
## Vector infection data

```{r mosquitodataread, echo=FALSE} 

wnv <- read.csv(mosqfile, stringsAsFactors=FALSE)
wnv$col_date <- as.Date(wnv$col_date, "%m/%d/%Y")
wnv$year <- as.numeric(format(wnv$col_date, "%Y"))

```


```{r mosquitodataprocess, echo=FALSE}
# convert district to factor
wnv$district <- simplifynames(wnv$district)
wnv$district <- factor(wnv$district)

# convert date to a date object
wnv$col_date <- as.Date(wnv$col_date, '%m/%d/%Y')

# figure out how many rows we start with
nrow1 <- nrow(wnv)

# create some variables we can use to filter
wnv$col_year <- as.numeric(format(wnv$col_date, "%Y"))
wnv$doy      <- as.numeric(format(wnv$col_date, "%j"))
wnv$weeknum  <- as.numeric(format(wnv$col_date, "%U"))
wnv$species <- NULL
wnv$district <- factor(wnv$district)
wnv$district <- droplevels(wnv$district)

# figure out which years we're modeling
minmosqyear <- min(wnv$year, na.rm=TRUE)

# get rid of those which don't have a result
wnv <- wnv[which(!is.na(wnv$wnv_result)),]
wnv <- wnv[which(!is.na(wnv$doy)),]

# delete anything before a certain day
wnv <- wnv[which(wnv$doy >= 100),]
# delete anything after a certain day
wnv <- wnv[which(wnv$doy <= 212),]

# after cleaning, how many do we have?
nrow2 <- nrow(wnv)
nrow3 <- nrow(wnv[wnv$year == maxmosqyear,])

tempdf <- wnv[wnv$year == maxmosqyear,]
tempdf <- tempdf[!is.na(tempdf$wnv_result),]
wnvdenominator <- nrow(tempdf)
wnvnumerator <- nrow(tempdf[tempdf$wnv_result == 1,])

numpos <- wnvnumerator
perpos <- 100*round(wnvnumerator/wnvdenominator, 3)
```

There are `r nrow2` samples in the vector testing database. For `r maxmosqyear`, there are `r nrow3` tested samples, with `r numpos` (`r perpos`%) positive. The estimated risk of human infection due to the early-season vector infection growth rate is shown below. Higher means that the pathogen is spreading more rapidly among vectors, and more human cases should be expected. The regions used for stratification are mapped below; districts are thought to share risk more closely with others in the same strata, although all districts share some level of risk with all other districts.  

```{r mosquitodataprocess2, fig.width=7, fig.height=3, echo=FALSE, include=FALSE}
# import district identifiers
strata <- read.csv(stratafile)
strata$district <- simplifynames(strata$district)
strata <- strata[c("district", "strata")]
wnv <- merge(x=wnv, y=strata,
             by.x="district",
             by.y="district",
             all.x=TRUE)

# figure out how many distinct districts we have left
districtlist           <- data.frame(district=unique(wnv$district))
distinctdistricts      <- length(unique(wnv$district))
districtlist$districtnum <- seq(from=1, to=distinctdistricts, by=1)
wnv <- merge(x=wnv, y=districtlist,
             by="district",
             all=TRUE)

# create a variable that at least has a little chance of being orthogonal to 1.
wnv$dminus <- wnv$doy - mean(wnv$doy, na.rm=TRUE)

# make sure all the observations have a stratum and year
wnv <- wnv[!is.na(wnv$strata),]
wnv <- wnv[!is.na(wnv$year),]

# run a random effect model on orthogonalized data
infectglm <- glmer(wnv_result ~ 1+dminus+
                    (0+1|year) +
                    (0+dminus|year) +
                    (0+1|strata:year) + 
                    (0+dminus|strata:year),
                  family=binomial(),
                  data=wnv)

wnv$est <- predict(infectglm, newdata=wnv, type="response")
write.csv(x=wnv, file=paste(mosqmatoutputdir,
                            "mosqmatrix.csv",
                            sep=""))

# predict random effects for all years
# dminus cannot be set to simply 0 - we have to make sure this isn't read in as a factor
randeffs <- expand.grid(strata=unique(wnv$strata),
                        year=minmosqyear:maxmosqyear,
                        dminus=0.00001)
randeffs <- randeffs[which(!is.na(randeffs$strata)),]
# if you don't allow new levels, the most recent year might not have an estimate
randeffs$mosqinfect <- predict(infectglm, newdata=randeffs, allow.new.levels = TRUE)
randeffs$stratayear <- paste(randeffs$strata, randeffs$year, sep=":")
randeffs$adjmosqinfect <- randeffs$mosqinfect - mean(randeffs$mosqinfect,na.rm=TRUE)
randeffs$stratum <- factor(randeffs$strata)
thisplot <- ggplot(randeffs) + geom_line(aes(x=year, y=adjmosqinfect, group=stratum, color=stratum)) +
  geom_abline(slope=0, intercept=0, linetype=2) +
  scale_x_continuous(breaks=minmosqyear:maxmosqyear) +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        panel.grid.minor=element_blank(),
        panel.grid.major.y=element_blank())+
  xlab("") + ylab("Relative risk due to\nvector infection growth rate")
```
```{r silentsavemosq, include=FALSE, echo=FALSE}
plot(thisplot)
ggsave(paste(graphicoutputdir,
             "vector infection rates.png",
             sep=""), thisplot)
```

```{r humandata, include=FALSE, echo=TRUE}

# import data
human <- read.csv(humandatafile)
begrow <- nrow(human)
human$chardate <- as.character(levels(human$date))[as.numeric(human$date)]
human$date <- as.Date(human$date, "%m/%d/%Y")
human$creationyear <- as.numeric(format(human$date, "%Y"))
human$creationmonth <- as.numeric(format(human$date, "%m"))
# simplify district names
human$district <- simplifynames(human$district)
human$district <- factor(human$district)
human$doy <- as.numeric(format(human$date, "%j"))

# retain only those in the right date range
hWNVminyear <- minmosqyear
hWNVmaxyear <- maxmosqyear-1
human <- human[which((human$creationyear >= hWNVminyear) & (human$creationyear <= hWNVmaxyear)),]

# retain only those in months 5-11
human <- human[which((human$creationmonth >= 5) & (human$creationmonth <= 11)),]

# create the full list of weeks
minobservedhumandate <- min(human$date)

# set up the data frame so that it ends at the maxdesiredhumandate 
filledweeks <- seq(from=maxdesiredhumandate, to=minobservedhumandate, by=-7)

fullcasemat <- expand.grid(sort(unique(human$district)), filledweeks)
names(fullcasemat) <- c("district", "weekstartdate")
head(fullcasemat)
fullcasemat$anycases   <- rep(0, nrow(fullcasemat))
fullcasemat$totalcases <- rep(0, nrow(fullcasemat))
fullcasemat$observed   <- 1*((fullcasemat$weekstartdate >= minobservedhumandate)&
                             (fullcasemat$weekstartdate <= maxobservedhumandate))
for (i in 1:nrow(fullcasemat)) {
  
  thisweekstartdate <- fullcasemat$weekstartdate[i]
  thisdistrict        <- fullcasemat$district[i]
  
  tempcases <- human[which(human$district == thisdistrict),]
  tempcases <- tempcases[which(tempcases$date >= thisweekstartdate),]
  tempcases <- tempcases[which(tempcases$date <= (thisweekstartdate + 6)),]
  
  fullcasemat$anycases[i] <- 1*(nrow(tempcases) > 0)
  if (nrow(tempcases) > 0) {
    
    fullcasemat$totalcases[i] <- nrow(tempcases)
    
  }
  
}
totcase <- sum(fullcasemat$totalcases, na.rm=TRUE)
anypos  <- sum(fullcasemat$anycases, na.rm=TRUE)

# figure out what percentage of cases we'll likely have seen before the start of this week
weekinquestionSundoy <- as.numeric(format(weekinquestionSun, "%j"))
fullcasemat$doy      <- as.numeric(format(fullcasemat$weekstartdate, "%j"))

tempdf <- fullcasemat[fullcasemat$doy < (weekinquestionSundoy+7),]
observedbefore <- sum(tempdf$totalcases, na.rm=TRUE)
observedtotal  <- sum(fullcasemat$totalcases, na.rm=TRUE)
weekinquestionreformat <- format(weekinquestionSun, "%m-%d")
observedfraction <- 100*round(observedbefore / observedtotal, 2)
```
```{r stratamap, include=FALSE, echo=FALSE}

district_shapes <- readShapePoly(districtshapefile)
# simplify name
district_shapes$district <- simplifynames(district_shapes$NAME)
crs(district_shapes) <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80     +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
projected_districts <- spTransform(district_shapes, crs("+proj=longlat +datum=WGS84 +no_defs"))
projected_districts@data$id = rownames(projected_districts@data)
projected_districts.df <- tidy(projected_districts)
projected_districts.df <- left_join(projected_districts.df, projected_districts@data, by="id")

stratamapcsv <- read.csv(stratafile)
stratamapcsv$district <- simplifynames(stratamapcsv$district)
stratamapcsv <- stratamapcsv[c("district", "strata")]

projected_districts.df <- left_join(projected_districts.df, stratamapcsv,
                                   by="district")

projected_districts.df$stratum <- factor(projected_districts.df$strata.x)
```
```{r stratmap2, include=TRUE, echo=FALSE, fig.width=7, fig.height=3}
thisplot <- ggplot(projected_districts.df) +
      aes(long,lat,fill=stratum,group=group,id=id,guides=FALSE) +
      geom_polygon() + xlab("") + ylab("") +
      geom_path(color="black") +
      theme(legend.position="bottom") +
      coord_map() + ggtitle("State stratification map") +
        theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
            axis.ticks = element_blank(), axis.text.x = element_blank(),
            axis.text.y = element_blank(), axis.title.x=element_blank(),
            axis.title.y = element_blank(),
            legend.position="right",
            legend.key.width=unit(1, "cm"),
            legend.key.height=unit(0.5,"cm"))
plot(thisplot)
```
```{r silentsave88791122823, include=FALSE, echo=FALSE}
ggsave(paste(graphicoutputdir,
             "strata.png",
             sep=""), thisplot)

```

The following graph shows the estimated growth of positive samples for every year (grey), with `r compyear1` (blue) and `r compyear2` (blue, dashed) selected for comparison, and estimates and observations for `r maxmosqyear` (red). The lines are modeled sample positive rates; the actual statewide positive sample rate for `r maxmosqyear` is shown here by grouping observations nearby in time.

```{r mosqbymonth, include=TRUE, echo=FALSE, fig.width=6, fig.height=3}
mosqmopreds <- expand.grid(strata=unique(wnv$strata),
                           year=minmosqyear:maxmosqyear,
                           doy=seq(from=min(wnv$doy, na.rm=TRUE),
                                   to  =max(wnv$doy, na.rm=TRUE), by=1))
mosqmopreds$dminus <- mosqmopreds$doy - mean(wnv$doy, na.rm=TRUE)
mosqmopreds$preds <- predict(infectglm, newdata=mosqmopreds, type="response", allow.new.levels=TRUE)

mosqmopreds <- group_by(mosqmopreds, dminus, year)
mosqmopreds <- dplyr::summarize(mosqmopreds,
                         preds=mean(preds, na.rm=TRUE),
                         doy=mean(doy, na.rm=TRUE))
thisyear1   <- mosqmopreds[mosqmopreds$year == maxmosqyear,]
comparison1 <- mosqmopreds[mosqmopreds$year == compyear1,]
comparison2 <- mosqmopreds[mosqmopreds$year == compyear2,]

thisyeardot <- wnv[wnv$year == maxmosqyear,]

if (sum(!is.na(thisyeardot$doy)) > 20) {

  thisyeardot$rounddoy <- cut2(thisyeardot$doy, g=6)
  thisyeardot <- group_by(thisyeardot, rounddoy)
  thisyeardot <- dplyr::summarize(thisyeardot,
                        meanpos = mean(wnv_result, na.rm=TRUE),
                        meandoy = mean(doy, na.rm=TRUE))
  thisplot <- ggplot(data=mosqmopreds) + geom_line(data=mosqmopreds, aes(x=doy, y=preds, group=year),
                                                   color="grey", alpha=0.5) +
    geom_line(data=comparison1, aes(x=doy, y=preds), color="blue") +
    geom_line(data=comparison2, aes(x=doy, y=preds), color="blue", linetype=2) +
    geom_line(data=thisyear1, aes(x=doy, y=preds), color="red") +
    geom_point(data=thisyeardot, aes(x=meandoy, y=meanpos), color="red") +
    xlab("day of the year") + ylab("Vector pool positive rate") +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"))
  
} else {
  
  thisplot <- ggplot(data=mosqmopreds) + geom_line(data=mosqmopreds, aes(x=doy, y=preds, group=year),
                                                   color="grey", alpha=0.5) +
    geom_line(data=comparison1, aes(x=doy, y=preds), color="blue") +
    geom_line(data=comparison2, aes(x=doy, y=preds), color="blue", linetype=2) +
    geom_line(data=thisyear1, aes(x=doy, y=preds), color="red") +
    xlab("day of the year") + ylab("Vector pool positive rate") +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"))
  
}


plot(thisplot)
```
```{r include=FALSE, echo=FALSE}
ggsave(paste(graphicoutputdir,
             "mosqinfectgrowthrates.png",
             sep=""), thisplot)
```


## Human data

The predictive model of human cases was calibrated using `r totcase` historical cases, not including any cases from `r maxmosqyear`. No cases from `r maxmosqyear` are used to make predictions; the estimates for this year are based solely on weather and vector data. Typically, `r observedfraction`% of a year's cases occur before the end of this week in any given year.

# Model results

## Statewide trends

The graphs below show observed statewide risk (black) and estimated risk (red) up to `r maxdesiredhumandatestr` `r maxdesiredhumandate`. Observed risk should be completely zero during the last year - these data are not used in the model, and will only be updated once final human case data are received at the end of the year.

```{r humandata2, echo=TRUE, warnings=FALSE, include=FALSE}
# select which variables we're going to use!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
dailyextr <- data.frame(dailyextr)
dailyextr$var1 <- dailyextr[,which(colnames(dailyextr) == var1name)]
dailyextr$var2 <- dailyextr[,which(colnames(dailyextr) == var2name)]
dailyextr <- dailyextr[c("district","date","var1","var2")]
dailyextr$district <- simplifynames(dailyextr$district)
# fill in missing and future climatological data
totaldailyextr <- expand.grid(district=unique(dailyextr$district),
                              date=seq(from=as.Date("2003-01-01", "%Y-%m-%d"),
                                       to=maxdesiredhumandate,
                                       by=1))
dailyextr <- merge(x=totaldailyextr, y=dailyextr,
                   by.x=c("district","date"),
                   by.y=c("district","date"),
                   all.x=TRUE)
dailyextr$doy <- as.numeric(format(dailyextr$date, "%j"))

dailyextr <- group_by(dailyextr, district, doy)
districtdoymean <- dplyr::summarize(dailyextr,
                           meanvar1=mean(var1, na.rm=TRUE),
                           meanvar2=mean(var2, na.rm=TRUE))
dailyextr <- ungroup(dailyextr)                           

dailyextr <- left_join(dailyextr, districtdoymean,
                   by=c("district","doy"))

dailyextr$var1[is.na(dailyextr$var1)] <- dailyextr$meanvar1[is.na(dailyextr$var1)]
dailyextr$var2[is.na(dailyextr$var2)] <- dailyextr$meanvar2[is.na(dailyextr$var2)]

# garbage collection
rm(districtdoymean)
dailyextr$meanvar1 <- NULL
dailyextr$meanvar2 <- NULL
gc()

# we need the districts to be in lower case to merge with the gridMET data
fullcasemat$district <- simplifynames(fullcasemat$district)

datalagger <- expand.grid(unique(fullcasemat$district),
                          unique(fullcasemat$weekstartdate),
                          seq(from=0, to=laglen-1, by=1))
names(datalagger) <- c("district","date","lag")
datalagger$laggeddate <- datalagger$date-datalagger$lag

datalagger <- left_join(datalagger, dailyextr,
                        by=c("district"="district",
                             "laggeddate"="date"))

# garbage collection
rm(dailyextr)
gc()

# pivot
mean1data <- dcast(datalagger, district + date ~ lag, value.var="var1")
names(mean1data) <- paste("var1_",names(mean1data),sep="")
head(mean1data)

mean2data <- dcast(datalagger, district + date ~ lag, value.var="var2")
names(mean2data) <- paste("var2_",names(mean2data),sep="")
head(mean2data)

# garbage collection
rm(datalagger)
gc()

# and put all this lagged info back into the total case matrix
fullcasemat <- merge(x=fullcasemat, y=mean1data,
                     by.x=c("district", "weekstartdate"),
                     by.y=c("var1_district","var1_date"),
                     all.x=TRUE)
fullcasemat <- merge(x=fullcasemat, y=mean2data,
                     by.x=c("district", "weekstartdate"),
                     by.y=c("var2_district","var2_date"),
                     all.x=TRUE)

# turn these into matrices
fullcasemat$var1 <- as.matrix(fullcasemat[,grep(x=colnames(fullcasemat),
                                                pattern="var1_",
                                                fixed=TRUE)])
fullcasemat$var2 <- as.matrix(fullcasemat[,grep(x=colnames(fullcasemat),
                                                pattern="var2_",
                                                fixed=TRUE)])
fullcasemat[,grep(x=colnames(fullcasemat), pattern="var1_", fixed=TRUE)] <- NULL
fullcasemat[,grep(x=colnames(fullcasemat), pattern="var2_", fixed=TRUE)] <- NULL

# import strata
fullcasemat <- merge(x=fullcasemat, y=strata,
                     by.x="district", by.y="district",
                     all.x=TRUE)

# stratify covariates
for (curstrata in unique(fullcasemat$strata)) {
  
  fullcasemat[c(paste("dl_var1_", curstrata, sep=""))] <- fullcasemat$var1 * (fullcasemat$strata == curstrata)
  fullcasemat[c(paste("dl_var2_", curstrata, sep=""))] <- fullcasemat$var2 * (fullcasemat$strata == curstrata)
  
}

# create lag matrix
fullcasemat$lag <- fullcasemat$var1
for (curcol in 1:dim(fullcasemat$lag)[2]) {
  
  fullcasemat$lag[,curcol] <- (curcol-1)

}

# garbage collection
rm(mean1data)
rm(mean2data)
gc()

# include the mosquito summary statistic
stratayearmosq <- randeffs
stratayearmosq <- stratayearmosq[c("stratayear","mosqinfect")]
names(stratayearmosq) <- c("stratayear","MIRsummarystat")

fullcasemat$year <- as.numeric(format(fullcasemat$weekstartdate, "%Y"))
fullcasemat$stratayear <- paste(fullcasemat$strata, fullcasemat$year, sep=":")
fullcasemat$year <- NULL

fullcasemat <- merge(x=fullcasemat, y=stratayearmosq,
                     by.x="stratayear", by.y="stratayear",
                     all.x=TRUE)

# calculate missing MIRsummarystat
stratayearmosq$strata <- substr(stratayearmosq$stratayear, 1, 3)
tempdf <- aggregate(stratayearmosq$MIRsummarystat,
                    by=list(stratayearmosq$strata),
                    FUN=mean,
                    na.rm=TRUE)
names(tempdf) <- c("strata","MIRsummarystat2")

fullcasemat <- merge(x=fullcasemat, y=tempdf,
                     by.x="strata", by.y="strata",
                     all.x=TRUE)
fullcasemat$MIRsummarystat[is.na(fullcasemat$MIRsummarystat)] <- fullcasemat$MIRsummarystat2[is.na(fullcasemat$MIRsummarystat)]

fullcasemat$strata <- NULL
fullcasemat$stratayear <- NULL
```

```{r humanreg, echo=FALSE, include=FALSE}

# get rid of districts that never have a positive case in the time period
fullcasemat$doy <- as.numeric(format(fullcasemat$weekstartdate, "%j"))

# make sure district is a factor
fullcasemat$district <- factor(fullcasemat$district)

myformula <- paste("s(lag, by=",
                   names(fullcasemat)[grep(x=names(fullcasemat),
                              pattern="dl_",
                              fixed=TRUE)],
                   ", bs='tp')",sep="", collapse=" + ")
myformula <- paste("anycases ~ 0 + MIRsummarystat + district", myformula, sep=" + ")
myform <- as.formula(myformula)

cl <- makeCluster(detectCores(logical=FALSE)-1)
firstreg  <- bam(formula=myform,
                 family=binomial(), data=fullcasemat,
                 subset=observed==1,
                 cluster=cl)
stopCluster(cl)

# show predictions
preds <- predict(firstreg, type="link", newdata=fullcasemat, se.fit=TRUE, newlevels=TRUE)
fullcasemat$est <- 1/(1 + exp(-preds$fit))

preds <- as.matrix(cbind(preds$fit, preds$se.fit))
preds <- data.frame(preds)
names(preds) <- c("est", "sd")
preds$upper <- preds$est + 1.96*preds$sd
preds$lower <- preds$est - 1.96*preds$sd
preds$est   <- preds$est

preds$upper <- 1 / (1 + exp(-preds$upper))
preds$lower <- 1 / (1 + exp(-preds$lower))
preds$est <- 1 / (1 + exp(-preds$est))

tempfull <- fullcasemat
tempfull$preds <- preds$est

tempfull$month <- as.numeric(format(tempfull$weekstartdate, "%m"))
tempfull <- tempfull[which(tempfull$month >= 6),]
tempfull <- tempfull[which(tempfull$month <= 10),]

```

```{r humanregplot, echo=FALSE}
tempdf <- data.frame(weekstartdate=fullcasemat$weekstartdate,
                     obs=fullcasemat$anycases,
                     est=preds$est)
tempdf <- group_by(tempdf, weekstartdate)
tempdf <- dplyr::summarize(tempdf,
                    obs=mean(obs, na.rm=TRUE),
                    est=mean(est, na.rm=TRUE))
tempdf <- ungroup(tempdf)
tempdf2 <- tempdf
tempdf2$week <- as.numeric(format(tempdf2$weekstartdate, "%U"))
lowe <- 12
hiwe <- 52-12

tempdf2 <- tempdf2[tempdf2$week >= lowe,]
tempdf2 <- tempdf2[tempdf2$week <= hiwe,]

tempdf2$newwe <- (tempdf2$week - lowe) / (hiwe - lowe + 1)
tempdf2$year <- as.numeric(format(tempdf2$weekstartdate, "%Y"))
tempdf2$newdate <- tempdf2$year + tempdf2$newwe
```

```{r fig.width=7, fig.height=3.5, echo=FALSE, warnings=FALSE}
thisplot <- ggplot(tempdf2) + geom_line(aes(x=newdate,
                               y=obs)) +
  geom_line(aes(x=newdate, y=est), color="red") +
  ggtitle("Statewide model predictions") +
  xlab("") + ylab("") +
  scale_x_continuous(breaks=seq(from=minmosqyear,to=(maxmosqyear+1),by=1),
                     labels=c(seq(from=minmosqyear, to=maxmosqyear, by=1), ""),
                     limits=c((minmosqyear+0.125),(maxmosqyear+1))) +
  ylab("Proportion of districts positive") +
  theme(panel.grid.minor.x=element_blank()) +
  theme(axis.text.x = element_text(hjust=-0.225))
plot(thisplot)
```
```{r silentsave8879823, include=FALSE, echo=FALSE}
ggsave(paste(graphicoutputdir,
             "predictions.png",
             sep=""), thisplot)

```
```{r curyearplot, include=TRUE, echo=FALSE, fig.width=7, fig.height=3.5}

tempdf <- fullcasemat
tempdf <- group_by(tempdf, weekstartdate)
tempdf <- dplyr::summarize(tempdf,
                    obs=mean(anycases, na.rm=TRUE),
                    est=mean(est, na.rm=TRUE))
tempdf$year <- as.numeric(format(tempdf$weekstartdate, "%Y"))
tempdf$month <- as.numeric(format(tempdf$weekstartdate, "%m"))

tempdf <- tempdf[tempdf$month >= 5,]
tempdf <- tempdf[tempdf$month <= 11,]

thisyear <- tempdf[tempdf$year == maxmosqyear,]
comparison1 <- tempdf[tempdf$year == compyear1,]
comparison2 <- tempdf[tempdf$year == compyear2,]

comparison1$weekstartdate <- comparison1$weekstartdate + (maxmosqyear - compyear1)*365
comparison2$weekstartdate <- comparison2$weekstartdate + (maxmosqyear - compyear2)*365

tempdf <- bind_rows(thisyear,
                    comparison1,
                    comparison2)

tempdf$est[tempdf$year == compyear1] <- tempdf$obs[tempdf$year == compyear1]
tempdf$est[tempdf$year == compyear2] <- tempdf$obs[tempdf$year == compyear2]
tempdf$year <- factor(tempdf$year)

tempdf2 <- thisyear[thisyear$weekstartdate >= weekinquestionSun,]
tempdf2 <- tempdf2[1,]
```
Predictions for `r maxmosqyear` are compared to observations in `r compyear1` and `r compyear2`. We expect `r round(66*tempdf2$est,1)` districts to have at least one human case between `r weekinquestionSunstr` `r weekinquestionSun` and `r weekinquestionSatstr` `r weekinquestionSat`.

```{r plotblah, include=TRUE, echo=FALSE, fig.width=7, fig.height=3.5}
thisplot <- ggplot(tempdf) + geom_line(data=tempdf, aes(x=weekstartdate,
                           y=est,
                           color=year,
                           linetype=year)) +
  geom_point(data=tempdf2, aes(x=weekstartdate, y=est), color="red", size=4)+
  xlab("") + ylab("") +
  scale_x_date(date_labels="%b", date_breaks="1 month",
               limits=c(as.Date(paste(maxmosqyear, "-05-01", sep=""), "%Y-%m-%d"),
                        as.Date(paste(maxmosqyear, "-11-15", sep=""), "%Y-%m-%d")),
               labels=c("May","Jun","Jul","Aug","Sep","Oct","Nov","")) +
  ylab("Proportion of districts positive") +
  theme(panel.grid.minor.x=element_blank()) +
  scale_linetype_manual(values=c(3,2,1)) +
  scale_color_manual(values=c("black","black","red"))
thisplot1 <- thisplot + ggtitle(paste("Estimates for ",
                maxmosqyear, " compared to observations in ", compyear1, " and ", compyear2,
                "\nwith week beginning ", weekinquestionSun, " highlighted", sep="")) +
  theme(axis.text.x = element_text(hjust=-1.25))
plot(thisplot1)
```
```{r silentsave12899234, include=FALSE, echo=FALSE}
thisplot2 <- thisplot + theme(legend.position="bottom") +
  theme(axis.text.x = element_text(hjust=-0.50))
ggsave(paste(graphicoutputdir,
             "estimates.png",
             sep=""), thisplot2,
       width=4,
       height=4)

```

```{r echo=FALSE}
write.csv(x=fullcasemat[c("district","weekstartdate",
                          "anycases","totalcases", "est")],
          file=paste(fullcasematoutputdir, "case matrix.csv", sep=""))
fullcasemat$year <- as.numeric(format(fullcasemat$weekstartdate, "%Y"))
thisyear <- fullcasemat[fullcasemat$year == maxmosqyear,]
previousyears <- fullcasemat[fullcasemat$year < maxmosqyear,]
sumtotal <- sum(previousyears$totalcases, na.rm=TRUE)
sumany   <- sum(previousyears$anycases, na.rm=TRUE)

multiplier <- round(sumtotal / sumany, 2)
positivesthisyear <- round(sum(thisyear$est, na.rm=TRUE), 1)
casesthisyear     <- round(positivesthisyear * multiplier, 1)
```

We expect that there will be `r positivesthisyear` positive district-weeks this year. If we assume that each positive district-week corresponds to `r multiplier` cases, which is an estimate based on historical data, then this yields a total of `r casesthisyear` cases estimated for `r maxmosqyear`.

\newpage
## Results for `r weekinquestionSun` to `r format(weekinquestionSat, "%m-%d")`

We visualize the raw estimated risk for `r weekinquestion` below. If a district is darkest blue, then we estimate that there should be no human cases reported for this district, during this week. If a district is brightest red, we are certain that there will be at least one human case reported for this district, during this week.

```{r shapefile, include=FALSE, echo=FALSE, warnings=FALSE}
district_shapes <- readShapePoly(districtshapefile)
district_shapes$NAME <- simplifynames(district_shapes$NAME)
crs(district_shapes) <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80     +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
projected_districts <- spTransform(district_shapes, crs("+proj=longlat +datum=WGS84 +no_defs"))
projected_districts@data$id = rownames(projected_districts@data)
projected_districts.df <- tidy(projected_districts)
projected_districts.df <- left_join(projected_districts.df, projected_districts@data, by="id")
projected_districts.df$district <- simplifynames(projected_districts.df$NAME)

# get this week's predictions
thisweek <- fullcasemat[fullcasemat$weekstartdate >= weekinquestion,]
thisweek <- thisweek[thisweek$weekstartdate == min(thisweek$weekstartdate, na.rm=TRUE),]
thisweek <- thisweek[c("district", "est")]

projected_districts.df <- left_join(projected_districts.df, thisweek,
                                   by="district")
projected_districts.df$est[is.na(projected_districts.df$est)] <- min(projected_districts.df$est, na.rm=TRUE)
```
```{r shapefile2, include=TRUE, echo=FALSE, fig.width=7, fig.height=3.5}
thisplot <- ggplot(projected_districts.df) +
      aes(long,lat,fill=est,group=group,id=id,guides=FALSE) +
      geom_polygon() + xlab("") + ylab("") +
      geom_path(color="black") +
      theme(legend.position="bottom") +
      scale_fill_distiller(palette = "Spectral", limits=c(0,1),
                                                 breaks = c(0,1),
                                                 labels = c("Will definitely not\nreport any cases",
                                                            "Will definitely\nreport some cases"),
                                                     name = "") +
      coord_map() + ggtitle(paste("Estimate for week beginning ",
                            weekinquestionSun,
                            sep="")) +
        theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
            axis.ticks = element_blank(), axis.text.x = element_blank(),
            axis.text.y = element_blank(), axis.title.x=element_blank(),
            axis.title.y = element_blank(),
            legend.position="bottom",
            legend.key.width=unit(1, "cm"),
            legend.key.height=unit(0.25,"cm"))
plot(thisplot)
```
```{r silentsave89712, include=FALSE, echo=FALSE}
ggsave(paste(graphicoutputdir,
             "map absolute.png",
             sep=""), thisplot)
```

This map indicates whether probabilities reported in the previous map are higher (red) than average, lower (blue) than average, or right about normal (yellow) compared to the same week in previous years.

```{r riskcalcs, include=TRUE, echo=FALSE, fig.width=7, fig.height=3.5}
thisweek <- fullcasemat[fullcasemat$weekstartdate >= weekinquestion,]
thisweek <- thisweek[thisweek$weekstartdate == min(thisweek$weekstartdate, na.rm=TRUE),]
thisweek <- thisweek[c("district", "est", "doy")]

approxdoy <- thisweek$doy[1]

fullcasemat$year <- as.numeric(format(fullcasemat$weekstartdate, "%Y"))

doypreds <- data.frame()
for (curyear in unique(fullcasemat$year)) {
  
  for (curdistrict in unique(fullcasemat$district)) {
    
    thisdf <- fullcasemat[fullcasemat$year == curyear,]
    thisdf <- thisdf[thisdf$district == curdistrict,]
    
    if(sum(!is.na(thisdf$est)) > 1) {
    
      tempdf <- data.frame(district = curdistrict,
                           year = curyear,
                           est = approx(x=thisdf$doy,
                                        y=thisdf$est,
                                        xout=approxdoy)$y)
      
      doypreds <- bind_rows(doypreds, tempdf)
      
    }
    
  }
  
}

doypreds2 <- data.frame()
for (curdistrict in unique(doypreds$district)) {

  tempdf <- doypreds[doypreds$district == curdistrict,]
  tempdf$percentile <- rank(tempdf$est, ties.method="random") / length(tempdf$est)
  
  doypreds2 <- bind_rows(doypreds2, tempdf)
  
}
doypreds <- doypreds2
rm(doypreds2)

riskalpha <- 0.25
doypreds$riskcategory <- " About average    "
doypreds$riskcategory[doypreds$percentile <= riskalpha/2] <- " Lower than usual    "
doypreds$riskcategory[doypreds$percentile >= 1-(riskalpha/2)] <- " Higher than usual    "
doypredscurrent <- doypreds[doypreds$year == as.numeric(format(weekinquestion, "%Y")),]
projected_districts.df <- left_join(projected_districts.df, doypredscurrent,
                                   by="district")
projected_districts.df$riskcategory[is.na(projected_districts.df$riskcategory)] <- " Not able to model    "
thisplotrisk <- ggplot(projected_districts.df) +
      aes(long,lat,fill=riskcategory,group=group,guides=FALSE) +
      geom_polygon() + 
      geom_path(color="black") +
      coord_map() + 
      ggtitle("") +
      theme(text=element_text(size=15)) +
      theme(plot.title=element_text(size=15),
            legend.key=element_rect(fill="white")) +
      theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
            axis.ticks = element_blank(), axis.text.x = element_blank(),
            axis.text.y = element_blank(), axis.title.x=element_blank(),
            axis.title.y = element_blank(),
            legend.position="bottom",
            legend.key.width=unit(1, "cm"),
            legend.key.height=unit(0.5,"cm")) +
      theme(legend.direction="vertical") +
      scale_fill_manual(name=paste("Risk for ", weekinquestionSun, " to ",
                                   format(weekinquestionSat, "%m-%d"), sep=""),
                        values=c(" Lower than usual    "="lightblue",
                                 " About average    "="yellow",
                                 " Higher than usual    "="red",
                                 " Not able to model    "="grey"))


plot(thisplotrisk)
```
```{r silentsave1245, include=FALSE, echo=FALSE}
ggsave(paste(graphicoutputdir,
             "map relative.png",
             sep=""), thisplotrisk,
       width=4,
       height=4)

sitegraphic <- grid.arrange(thisplotrisk +
                            theme(legend.position="right",
                                  legend.direction="vertical") +
                            scale_fill_manual(name="Risk this week",
                            values=c(" Lower than usual    "="lightblue",
                                     " About average    "="yellow",
                                     " Higher than usual    "="red",
                                     " Not able to model    "="grey")) +
                            #ylab("hidden axis") +
                            #theme(axis.title.y = element_text(colour = "white")) +
                            theme(legend.key.width=unit(0.25, "cm"),
                            legend.key.height=unit(0.50,"cm"),
                            text=element_text(size=10)) + 
                            theme(plot.margin = unit(c(0,0,0,0.4), "cm")),
  
                            # t r b l
                            
                            thisplot2 +
                            theme(legend.position="right") + ylab("") +
                            theme(axis.text.x=element_text(family="mono"))+
                            theme(panel.grid.minor=element_blank(),
                                  axis.text.y=element_blank(),
                                  axis.ticks=element_blank())+
                            theme(axis.text.x = element_text(hjust=-0.35)) +
                            theme(plot.margin = unit(c(0.1,0.2,0.1,-0.1), "cm")), nrow=2) 

ggsave(paste(graphicoutputdir,
             "site graphic.png",
             sep=""), sitegraphic,
       width=4,
       height=4)

```




